---
title: "PCA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Principal components analysis

## Lecture

Before starting, you should watch the lecture by François Husson on a practical presentation of PCA. [french](https://www.youtube.com/playlist?list=PLnZgp6epRBbRn3FeMdaQgVsFh9Kl0fjqX),[english](https://www.youtube.com/playlist?list=PLnZgp6epRBbQuZSHJ5hBzxdDZNzBFKBoB)

Principal component analysis (PCA) summarizes a data table where the observations are described by (continuous) quantitative variables. PCA is used
to study the similarities between individuals from the point of view of all of the variables and identifies individuals profiles. It
is also used to study the linear relationships between variables (based on correlation coefficients). The two studies are linked, which means that the individuals or groups can be characterized by the variables and the links between variables illustrated
from typical individuals. 
More details are in the lecture slides.

To perform PCA, we use the PCA function from the FactoMineR package. A [website](http://factominer.free.fr) is dedicated to this package. 
The code associated to the lecture slides is the following:

```{r eval = FALSE}
library(FactoMineR)
#Expert <- read.table("http://factominer.free.fr/course/doc/data_PCA_ExpertWine.csv",
#header = TRUE, sep = ";", row.names = 1)
Expert <- read.table("Data/Expert_wine.csv", header = TRUE, sep = ";", row.names = 1)
res.pca <- PCA(Expert, scale = T, quanti.sup = 29:30, quali.sup = 1)
summary(res.pca)
barplot(res.pca$eig[,1], main = "Eigenvalues", names.arg = 1:nrow(res.pca$eig))
plot.PCA(res.pca, habillage = 1)
res.pca$ind$coord; res.pca$ind$cos2; res.pca$ind$contrib
plot.PCA(res.pca, axes = c(3, 4), habillage = 1)
dimdesc(res.pca)
plotellipses(res.pca, 1)
write.infile(res.pca, file = "my_FactoMineR_results.csv")
```

A Shiny interface has been developed for FactoMineR, you should install the Factoshiny package, it may give you an idea of what interfaces you should produce to facilitate the use of the methods you develop.

```{r eval = FALSE}
library(Factoshiny)
PCAshiny(Expert)
```

A very useful package is  FactoInvestigate. It generates an automatic report and interpretation of FactoMineR principal component analyses. The main function provided by the package is the function Investigate(), which can be used to create either a Word, PDF or a HTML report.

```{r eval = FALSE}
library("FactoInvestigate")
# Principal component analyses
Investigate(res.pca, file = "PCA.Rmd", document = "pdf_document", 
            parallel = TRUE)
```

The aim of this package is also to encourage you going further in the interpretation!

Finally, we should mention that there exists a package that has embedded FactoMineR within ggplot2!

```{r eval = FALSE}
library(factoextra)
```

It is also possible to use simultaneously clustering (with k-means and AHC) and PCA to better analyse the data. For instance, we could apply clustering on the principal components for instance. 

```{r eval = FALSE}
HCPC(res.pca)
```

References on principal component methods and on the FactoMineR package:
Exploratory Multivariate Analysis by Example
using R, Husson, Lê, Pagès (2017), Chapman & Hall
Multiple Factor Analysis by Example using R, Pagès (2015), CRC Press.

